Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-10 rebuttal.tex

\comment{p1, c1, l29-37: I would dispute that until recently the
  cheminformatics techniques have been closely guarded
  secrets. J. Chem. Doc. was founded in 1961 and other techniques even
  older, such as the Wiener index (1947), Wiswesser Line Notation
  (1949), etc. Not to mention the actual foundations of what we call
  cheminformatics back to the atomistic theory of the 19th century.}
\response{We have the updated the text on page 1 to attenuate this
  statement, stressing on the fact that much of the data in
  cheminformatics is proprietary, whereas only some of the techniques (such
  as SMILES canonicalization) are. Regarding the statement that atomistic
  theory laid the foundations of cheminformatics - this is certainly
  true in a very broad sense. For that matter the atomistic theory
  laid the ground for computational chemistry in general. In the
  supplemental history document, we make a distinction between
  computational chemistry and cheminformatics}

\comment{p1, c1, l32: I think the phrase miracle molecule could be
  more usefully replaced with new drugs or new small molecule
  therapeutics.}
\response{The text has been updated to use the term ``therapeutic
  molecule''}

\comment{P4, c2, l17: citation needed on the size of chemistry space.}
\response{The appropriate reference has been added}

\comment{P4, c2, l23: the isomorphism problem has already been
  mentioned previously in this manuscript.}
\response{The text has been updated and simplified}

@@

\comment{P5, c1, l29: the normal phrase used to describe this concept
  is the similar property principle. The authors should also provide a
  reference.}
\response{This portion of the text has been restructured to refer to
  the similarity property principle and also include the relevant
  reference}

\comment{P5, c1, l40: QSAR should be referenced to Hansch et al. Also
  perhaps some discussion on what the authors mean by referring to
  these approaches as traditional.}
\response{We have updated text to include references to Hansch and
  Free \& Wilson. Regarding the use of ``traditional'', we believe the
  text explicitly explains why - the fact that QSAR as originaly
  defined only considers ligand features. However, we have rephrased it to use
  ``traditionally'' since one can argue that methods such as docking
  and pharmacophores are also QSAR methods, but consider both ligand
  and receptor.}

\comment{P5, c1, l41: I would dispute the assertion that these methods
  "ignore reception interactions" since the aim is to identify a
  correlation of biological response (e.g. pIC50) with chemical
  structure. The biological response is an explicit measurement of
  receptor interactions on the protein.}  
\response{We have updated
  the text to explicitly note that QSAR models do not usually take
  into account \emph{receptor features}. However we note that while
  receptor information is implicit in the $IC_{50}$, the value also
  includes other non-receptor related features such as
  permeability. Furthermore, lack of receptor information in QSAR
  models has been noted as the origin of activity cliffs (Guha and Van
  Drie, \textit{J.~Chem.~Inf.~Model.}, \textbf{2008}, \textit{48},
  1716--1728), thus supporting the statement that traditional QSAR
  models do indeed miss important information on receptor-ligand
  interactions.}

\comment{P5, c2, l22: I would mention naive Bayesian classifiers as well as
this is perhaps the most widely applied method in the field.}
\response{We have included the Na\"{i}ve Bayes as well}

### 2011-11-10 paper.tex
more review responses

The major difficulty with enumeration (apart from the size of the
search space, which is thought to be in the order of $10^{60}$
molecules for 30 heavy atoms \cite{Bohacek:1996ve}) is that the
\emph{in silico} representation of molecules, the so-called molecular
graphs, are labeled objects (i.e. labeled graphs), while the atoms of
a molecule are of course not uniquely labeled. The mathematical
concept to tackle this problem is to consider orbits of labeled
molecular graphs under the operation of the symmetric group. However,
when enumerating molecular structures, one has thus to solve the
so--called isomorphism problem, where molecules that seem different
are actually equivalent (i.e., their atoms can be mapped on top of
each other by keeping the connectivity between atoms intact).  The
good news is that isomorphism can be solved in polynomial time for
molecular graphs due to the fact that molecules belong to a restricted
class of graphs known as \emph{bounded valence graphs}; however, that
same restriction prevents applying to a molecular graph the polynomial
delay enumeration results obtained for general graphs (Chapter 8
in~\cite{faulon2010}). While the complexity of molecular graph
enumeration remains an open problem it has nonetheless been shown that
molecules can be sampled efficiently~\cite{goldberg1999}. Sampling
procedures based on the Metropolis or the Genetic Algorithms have been
developed and used to elucidate natural compounds from NMR data.

>>>>

The major difficulty with enumeration (apart from the size of the
search space, which is thought to be in the order of $10^{60}$
molecules for 30 heavy atoms \cite{Bohacek:1996ve}) is that the
\emph{in silico} representation of molecules, the so-called molecular
graphs, are labeled objects (i.e. labeled graphs), while the atoms of
a molecule are of course not uniquely labeled. The mathematical
concept to tackle this problem is to consider orbits of labeled
molecular graphs under the operation of the symmetric group. While the
complexity of molecular graph enumeration remains an open problem it
has nonetheless been shown that molecules can be sampled
efficiently \cite{goldberg1999}. Sampling procedures based on the
Metropolis or the Genetic Algorithms have been developed and used to
elucidate natural compounds from NMR data.
